---
title: "MPH Capstone"
author: "Brandon Leggins"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
---

# Data Cleaning
```{r }

#load packages
library(haven)
library(dplyr)
library(tidyverse)


#read stata file 

chis2024 <- read_dta("ADULT.dta")



#data cleaning
newchis <- chis2024 %>%
 
  mutate(
   
      agegroup = fct_collapse(factor(srage_p1),
      "18-25" = c("18"),
      "26-39" = c("26", "30", "35"),
      "40-64" = c("40", "45", "50", "55", "60"),
      "65+"   = c("65", "70", "75", "80", "85")
    ), 
    
    # Recode Stable House: (Using a binary logic)
    instability = case_when(
      stablehouse == 1 ~ "no", # Very/Fairly/Somewhat stable
      stablehouse == 2 ~ "yes",  # Fairly and very unstable
      TRUE ~ NA_character_
    ),
    
    # Detailed Housing Stability (AM183)
    noncollaspedhousing = factor(am183,
                                 levels = c(1, 2, 3, 4, 5),
                                 labels = c("very stable", "fairly stable", "somewhat stable", 
                                            "fairly unstable", "Very unstable")),

    # Recode variable on worry about rent/mortgage (AM184)
    unaffordability = case_when( 
                          am184 %in% c(1,2) ~ "yes",
                          am184 %in% c(3,4) ~ "no", #not housing insecure is the ref            
                        TRUE ~ NA_character_
    ),
    
    # Create Housing Insecurity variable. If they said yes for either unaffordable or instability, then housing insecurity = yes
    housing_insecure = case_when(
      unaffordability == "yes" | instability == "yes" ~ "yes",
      unaffordability == "no" & instability == "no"   ~ "no",
      TRUE ~ NA_character_
    ),
  
    # Demographics
    gender = factor(srsex, levels = c(1, 2), labels = c("Male", "Female")),
    
    race = factor(case_when(
  ombsrr_p1 %in% c(4, 6) ~ "Other",
  ombsrr_p1 == 1 ~ "Hispanic",
  ombsrr_p1 == 2 ~ "White",
  ombsrr_p1 == 3 ~ "Black",
  ombsrr_p1 == 5 ~ "Asian"), levels = c("Asian", "Black", "Hispanic", 
               "Other","White")),
    
    education = factor(sreduc, 
                       levels = 1:4, 
                       labels = c("< HS", "HS education", "Some college", "College degree and above")),
    
    urban = factor(ur_clrt2, levels = c(1, 2), labels = c("Urban", "Rural")),
    
    poverty = factor(povll, levels = c(1:4),
                            labels = c("0-99%", "100-199%", "200-299%", "300% and above")),
    
    
    marital = case_when(
  marit2 %in% c(1, 2) ~ "partnered",
  marit2 %in% c(3, 4) ~ "unpartnered"
) %>% factor(),
   
  bmi = factor(rbmi, levels = c(1:4),
                      labels = c("underweight", "normal", "overweight", "obese")),

  physical_activity = factor(ac212, levels = c(1,2),
                                    labels = c("yes", "no")),



    # Simple Yes/No for Diabetes 
    diabetes = factor(ab22v2, levels = c(1, 2), labels = c("yes", "no")),
   
    
    housingdiscrimination = factor(am189, levels = c(1,2), labels  = c("yes", "no")),
    
    #**diabetes management
    #doctor said A1c is greater than 9%?. only those who responded yes to ever being told they have diabetes sees this question
   uncontrolleddiab = factor(ifelse(ab150 %in% c(-1, 3), NA, ab150), #-1 are participants who did not get the question because they never had diabetes. 3 are participants who are unsure
                           levels = c(1,2), 
                           labels = c("no", "yes")),
   
   # number of times they checked a1c in last year
      a1c_check = factor( case_when(
    ab27_p1 %in% c(0, 1) ~ "< 1 times",
    ab27_p1 %in% c(2, 3, 4, 5)     ~ ">= 2 times",  
    ab27_p1 < 0             ~  NA_character_, 
    TRUE                   ~  NA_character_  # Everything else becomes NA
  ), 
  levels = c("< 1 times", ">= 2 times")),
  
  
  #doctors provided developed diabetes care plan
    diabetesplan = factor(ifelse(ab112 %in% c(-1), NA, ab112), #-1 are participants who did not get the question because they never had diabetes. 
                           levels = c(1,2), 
                           labels = c("yes", "no")),
   
  
    #Currently taking insulin or pills to manage diabetes. 
    ##only those who responded yes to ever being diagnoses with diabetes sees this
   
   diabetesmeds = factor(
  case_when(
    diamed %in% c(1, 2, 3) ~ "yes meds",
    diamed == 4            ~ "no meds",  
    diamed < 0             ~  NA_character_, 
    TRUE                   ~  NA_character_  # Everything else becomes NA
  ), 
  levels = c("yes meds", "no meds")),

  generalhealth = factor(ab1, 
                  levels = c(1:5), 
                  labels = c("Excellent", "Very good", "Good", 
                             "Fair", "Poor")),

  hypertension = factor(ab29v2, 
                  levels = c(1:3), 
                  labels = c("Yes", "No", "Borderline hypertension")),

  heartdisease = factor(ab34, 
                  levels = c(1,2), 
                  labels = c("Yes", "No")),

   cholesterol = factor(ab154, 
                  levels = c(1:3), 
                  labels = c("Yes", "No", "Dont know")),

    )
```
## Frequency tables
```{r check}
# Checking my work to see if variables are recoded correctly 


sum(is.na(newchis$housing_insecure)) # checks how many NAs
table(newchis$diabetes, newchis$ab22v2) #outcome of diabetes check
table(newchis$diabetes)

table(newchis$race, newchis$ombsrr_p1) #race



```


# Weighted table 1
```{r}
#unweighted table
library(gtsummary)
library(srvyr)
library(survey)




#creating survey weights for weighted table 1
##removed type, scale, and mse options because I just need frequencies not 95%CI and variance
 chis_design1 <- svydesign(
  data    = newchis,
  weights = ~rakedw0,
  ids     = ~1          
)

 
 
#weighted table 1, with only percentages no counts
tbl1_weighted<- chis_design1 %>%
  tbl_svysummary(
    include = c(gender, agegroup, race, education, urban, poverty, marital, diabetes, bmi, physical_activity, generalhealth, hypertension, heartdisease),
    by = housing_insecure,
    type = list(gender ~ "dichotomous",
                urban ~ "dichotomous",
                marital ~ "dichotomous"),
    value = list(gender ~ "Male",
                 urban ~ "Urban",
                 marital ~ "partnered"),
    label = list(gender ~ "Male",
                 urban ~ "Urban Residence",
                 agegroup ~ "Age Group",
                 race ~ "Race/Ethnicity",
                 education ~ "Education",
                 poverty ~ "Poverty Level",
                 marital ~ "Married/Living with partner",
                 diabetes ~ "Diabetes yes",
                bmi ~ "BMI",
                physical_activity ~ "Physical Activity yes",
                generalhealth ~ "Self reported health",
                hypertension ~ "Hypertension",
                heartdisease ~ "Heart disease"
                ),
    statistic = list(all_categorical() ~ "{p}"), # Added % symbol for clarity
    digits = list(all_categorical() ~ 1)        # This ensures 1 decimal place (e.g., 10.1)
  ) %>%
  add_overall() %>%
  # FLIP THE COLUMN ORDER HERE: Move 'stat_2' (yes) before 'stat_1' (no)
  modify_table_body(~ .x %>% relocate(stat_2, .before = stat_1)) %>%
  modify_spanning_header(all_stat_cols() ~ "**Housing Insecurity Status**") %>%
  bold_labels() 

tbl1_weighted

#save table as word doc
library(flextable)
tbl1_weighted %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "table1weighted.docx")


```




# Preparing dataset for weighted regression models 
```{r analysis}
library(survey)
library(gtsummary)
#assigning reference levels
newchis <- newchis %>%
  mutate(
    unaffordability = relevel(as.factor(unaffordability), ref = "no"),
    instability = relevel(as.factor(instability), ref = "no"),
    housing_insecure= relevel(as.factor(housing_insecure), ref = "no"),
    agegroup           = relevel(as.factor(agegroup),           ref = "65+"),
    gender             = relevel(as.factor(gender),             ref = "Male"),
    poverty            = relevel(as.factor(poverty),            ref = "300% and above"),
    education         = relevel(as.factor(education), ref = "College degree and above"),
    housingdiscrimination     = relevel(as.factor(housingdiscrimination), ref = "no"),
    marital = relevel(as.factor(marital), ref = "partnered"),
    bmi = relevel(as.factor(bmi), ref = "normal"),
    physical_activity = relevel(as.factor(physical_activity), ref = "no"),
    
    # 2. Outcomes for Models (Numeric 0/1 -). #poisson model needs 0 and 1 variable
    diabetes_num = ifelse(diabetes == "yes", 1, 0),
    uncontrolled_num = ifelse(uncontrolleddiab == "yes", 1, 0),
    diabetesmeds_num = ifelse(diabetesmeds == "yes meds", 1, 0),
    diabetesplan_num = ifelse(diabetesplan == "no", 1, 0),
    a1c_check_num = ifelse(a1c_check == "< 1 times", 1, 0),
    
  )


#CHIS suggested code for survey weights
chis_design <- svrepdesign(data = newchis , weights = ~ rakedw0 , 
                           repweights = "rakedw[1-80]" , 
                           type = "other" , scale = 1 ,
                           rscales = 1 , mse = TRUE)
```

# Bar graph of diabetes prevelance by housing insecurity status and race 
```{r figure1}

library(ggplot2)



prop_table <- svyby(~diabetes_num, ~race + housing_insecure, 
                    design = chis_design, svymean, na.rm = TRUE) %>%
  mutate(
    race = factor(race, levels = c("Asian", "Black", "Hispanic", "Other", "White")),
    # Calculate 95% CI using the standard formula: mean +/- (1.96 * SE)
    lower_ci = diabetes_num - (1.96 * se),
    upper_ci = diabetes_num + (1.96 * se)
  )

# 2. Plot with error bars
figure1 <- ggplot(prop_table, aes(x = race, y = diabetes_num, fill = housing_insecure)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  # Add the 95% CI error bars
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    position = position_dodge(0.9), # Must match the geom_bar dodge width exactly
    width = 0.25,                    # Width of the horizontal whisker caps
    color = "black"
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(
       y = "Prevalence (%)",
       x = "Race/Ethnicity",
       fill = "Housing Insecurity") +
  theme_minimal()

figure1 
#save to copy to word 
ggsave(
  filename = "diabetes_prevalence_hd.png", 
  plot = figure1, 
  width = 8,       # Dimensions in inches, perfect for a standard Word margin
  height = 5, 
  dpi = 300        # This sets the high-definition resolution
)
```

# Bar graph of housing insecurity prevelance by race
```{r figure2}
#weighted housing insecurity prevelance by race
prop_table_housing <- svyby(~housing_insecure, ~race, 
                            design = chis_design, svymean, na.rm = TRUE) %>%
  as.data.frame() %>%  
  mutate(
    race = factor(race, levels = c("Asian", "Black", "Hispanic", "Other", "White")),
    # Using 'se2' because R drops the text labels for standard errors in factors
    lower_ci = housing_insecureyes - (1.96 * se2),
    upper_ci = housing_insecureyes + (1.96 * se2)
  )

figure2 <- ggplot(prop_table_housing, aes(x = race, y = housing_insecureyes)) +
  geom_bar(stat = "identity", fill = "#00BFC4", show.legend = FALSE) +  
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    width = 0.15,      
    color = "black"
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(
       y = "Housing Insecurity Prevalence (%)",
       x = "Race/Ethnicity") +
  theme_minimal()

figure2
#save to copy to word 
ggsave(
  filename = "housing_insecure_prevalence_hd.png", 
  plot = figure2, 
  width = 8,       # Dimensions in inches, perfect for a standard Word margin
  height = 5, 
  dpi = 300        # This sets the high-definition resolution
)
```
## Weighted ORs
Diabetes is not rare in this sample (>10%) so we do not want to intepret odds ratio because it may overestimate the effect. But we should still check the ORs
```{r}

#weighted odds ratio for diabetes

library(survey)
library(ggplot2)



#Odds ratio
modelOR<- svyglm(diabetes ~  unaffordability + 
                agegroup + race
                + gender + poverty + education,
                design = chis_design,
                family=quasibinomial(),
                   rescale=TRUE)

modelOR %>%
  tbl_regression(exponentiate = TRUE,
                 include = unaffordability) %>%  # only show housing insecurity
  bold_labels() %>%
  modify_column_hide(p.value) %>% #hide p value 
  modify_caption("**OR model 1**")


```
 
## Aim 1: Weighted prevalence ratios (PR)
Association of housing insecurity and diabetes, adjusting for age, gender, poverty, and education
```{r poisson}


#unadjusted model 1 

model1<- svyglm(diabetes_num ~ housing_insecure,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)

model1_tbl <- model1 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = housing_insecure,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(housing_insecure ~ "Housing Insecurity") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 1 unadjusted: Prevalence Ratios for Diabetes**")  %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


#model 2 adjusted model
model2<- svyglm(diabetes_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


model2_tbl <- model2 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = housing_insecure,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(housing_insecure ~ "Housing Insecurity") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 2: Prevalence Ratios for Diabetes**")  %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


#model 3: bmi and physical activity added to model 2
model3<- svyglm(diabetes_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital+bmi+physical_activity,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


model3_tbl <- model3 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = housing_insecure,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(housing_insecure ~ "Housing Insecurity") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 3: Prevalence Ratios for Diabetes**")  %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")



#merge models 1-3
#merge adjusted and unadjusted tables
aim1_tbl<- tbl_merge(
  tbls = list(model1_tbl, model2_tbl, model3_tbl),
  tab_spanner = c("**Unadjusted**", "**Adjusted**", "**Fully Adjusted**") 
) %>%
  modify_caption("**Model 1-3: Prevalence Ratios for Diabetes**")

aim1_tbl

#convert aim 1 table to word
aim1_tbl %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "aim1_table.docx")

```


## Aim 2: stratum specific PRs for race/ethnicity
```{r}


# 1. Get the list of all your race groups
race_levels <- levels(chis_design$variables$race)

# 2. Loop to build the 3-column table for EVERY individual race group
race_tables_list <- lapply(race_levels, function(r) {
  
  # Filter data for the current race group
  race_subset <- subset(chis_design, race == r)
  
  # Create the customized text label for this specific race strata
  strata_label <- paste0("Housing Insecurity (", r, " strata)")
  
  # Step B: Run the Unadjusted Model
  m1_race <- svyglm(diabetes_num ~ housing_insecure, 
                    design = race_subset, 
                    family = poisson(link = "log"), rescale = TRUE) %>%
    tbl_regression(exponentiate = TRUE, 
                   include = housing_insecure, 
                   label = list(housing_insecure ~ strata_label)) %>% # <-- UPDATED LABEL
    bold_labels() %>% modify_column_hide(p.value)
  
  # Step C: Run the Adjusted Model
  m2_race <- svyglm(diabetes_num ~ housing_insecure + agegroup + gender + poverty + education + marital, 
                    design = race_subset, 
                    family = poisson(link = "log"), rescale = TRUE) %>%
    tbl_regression(exponentiate = TRUE, 
                   include = housing_insecure, 
                   label = list(housing_insecure ~ strata_label)) %>% # <-- UPDATED LABEL
    bold_labels() %>% modify_column_hide(p.value)
  
  # Step D: Run the Adjusted + Health Model
  m3_race <- svyglm(diabetes_num ~ housing_insecure + agegroup + gender + poverty + education + marital + bmi + physical_activity, 
                    design = race_subset, 
                    family = poisson(link = "log"), rescale = TRUE) %>%
    tbl_regression(exponentiate = TRUE, 
                   include = housing_insecure, 
                   label = list(housing_insecure ~ strata_label)) %>% # <-- UPDATED LABEL
    bold_labels() %>% modify_column_hide(p.value)
  
  # Step E: Merge these 3 specific models side-by-side
  tbl_merge(
    tbls = list(m1_race, m2_race, m3_race), 
    tab_spanner = c("**Unadjusted**", "**Adjusted**", "**Fully adjusted**")
  )
})

# 3. Stack all of those mini tables vertically on top of each other
aim2_tbl_long <- tbl_stack(
  tbls = race_tables_list
  # Removed group_header entirely so it doesn't create the extra header rows
) %>%
  modify_caption("**Table: Race Stratified Prevalence Ratios for Diabetes by Housing Insecurity**") %>%
  modify_table_styling(
    columns = everything(),
    footnote = "PR = Prevalence Ratio; CI = Confidence Interval"
  )

aim2_tbl_long

#convert aim 2 table to word
aim2_tbl_long %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "aim2_table.docx")



```

### Forest Plot of PRs stratified by race
```{r forest}

library(broom)

library(ggplot2)

# 1. Programmatically extract the estimates and CIs from the models
plot_data <- lapply(race_levels, function(r) {
  model <- svyglm(diabetes_num ~ housing_insecure + agegroup + gender + poverty + education,
         design = subset(chis_design, race == r),
         family = poisson(link = "log"),
         rescale = TRUE)
  
  tidy_model <- tidy(model, exponentiate = TRUE, conf.int = TRUE)
  
  tidy_model %>%
    filter(grepl("housing_insecure", term)) %>% 
    mutate(race_stratum = r)
    
}) %>% 
  bind_rows() %>%
  mutate(race_stratum = fct_relevel(race_stratum, 
                                    "White", "Other", "Hispanic", "Black", "Asian"))

figure3 <- ggplot(plot_data, aes(x = estimate, y = race_stratum)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", linewidth = 0.8) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, 
                color = "darkblue", linewidth = 0.8, orientation = "y") +
  geom_point(color = "darkblue", size = 3.5) +
  labs(
    subtitle = "Association between housing insecurity and diabetes by race",
    x = "Prevalence Ratio (95% CI)",
    y = "Race/Ethnicity"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold")
  )

figure2

#save to copy to word 
  ggsave(
  filename = "figure3.png", 
  plot = figure3, 
  width = 8,       # Dimensions in inches, perfect for a standard Word margin
  height = 5, 
  dpi = 300        # This sets the high-definition resolution
)
```

## Aim 3 analysis: Diabetes Management
three seperate regressions with housing insecurity as the exposure and three unique outcomes: uncontrolled diabetes, clinical care, and diabetes monitoring. 

### Uncontrolled diabetes
```{r explore}

#unadjusted model 
uncontroldiab_model1 <- svyglm(uncontrolled_num ~ housing_insecure,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


uncontroldiab_model1_tbl<-uncontroldiab_model1 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models uncontrolled diabetes [yes] **")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")
 

#adjusted model
uncontroldiab_model2<- svyglm(uncontrolled_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


uncontroldiab_model2_tbl<-uncontroldiab_model2 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models uncontrolled diabetes [yes] **")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


#fully adjusted model
uncontroldiab_model3<- svyglm(uncontrolled_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital+bmi+physical_activity,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


uncontroldiab_model3_tbl<-uncontroldiab_model3 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models uncontrolled diabetes [yes] **")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


tbl_uncontrolled<- tbl_merge(
  tbls = list(uncontroldiab_model1_tbl, uncontroldiab_model2_tbl, uncontroldiab_model3_tbl),
  tab_spanner = c("**Unadjusted**", "**Adjusted**", "**Fully Adjusted**")
)

```

### Diabetes monitoring
```{r a1c}

#unadjusted model
diabmonitor_model1<- svyglm(a1c_check_num ~ housing_insecure,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


diabmonitor_model1_tbl<-diabmonitor_model1 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models less than 2 a1c checks a year **")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")

# Adjusted model
diabmonitor_model2<- svyglm(a1c_check_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


diabmonitor_model2_tbl<-diabmonitor_model2 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models less than two a1c checks a year **")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


# Fully Adjusted model
diabmonitor_model3<- svyglm(a1c_check_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital+bmi+physical_activity,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


diabmonitor_model3_tbl<-diabmonitor_model3 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models less than two a1c checks a year **")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")



tbl_monitoring<- tbl_merge(
  tbls = list(diabmonitor_model1_tbl, diabmonitor_model2_tbl, diabmonitor_model3_tbl),
  tab_spanner = c("**Unadjusted**", "**Adjusted**", "**Fully Adjusted**")
)


```

### Diabetes Care Plan
```{r plan}
#diabetes plan

#unadjusted
diabplan_model1<- svyglm(diabetesplan_num ~ housing_insecure,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


diabplan_model1_tbl<- diabplan_model1 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models not having a diabetes care plan**")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")

#adjusted
diabplan_model2<- svyglm(diabetesplan_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


diabplan_model2_tbl<-diabplan_model2 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models not having a diabetes care plan**")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")

#fully adjusted
diabplan_model3<- svyglm(diabetesplan_num ~ housing_insecure+agegroup+race+gender+poverty+education+marital+bmi+physical_activity,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


diabplan_model3_tbl<-diabplan_model3 %>%
  tbl_regression(exponentiate = TRUE, 
          include = housing_insecure,
          label = list(housing_insecure ~ "Housing Insecurity")) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  modify_caption("**models not having a diabetes care plan**")  %>%
  modify_header(estimate ~ "**PR**") %>% 
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


tbl_care<- tbl_merge(
  tbls = list(diabplan_model1_tbl, diabplan_model2_tbl, diabplan_model3_tbl),
  tab_spanner = c("**Unadjusted**", "**Adjusted**", "**Fully Adjusted**")
)

```

# Table 3: Diabetes management 
```{r tbl}
# stack all three outcomes into one table
aim3_tbl<-tbl_stack(
  tbls = list(tbl_care, tbl_monitoring, tbl_uncontrolled),
  group_header = c("No Diabetes Care Plan", "Poor Diabetes Monitoring", "Uncontrolled Diabetes")
) %>%
  modify_caption("**Diabetes Management**")

aim3_tbl


#convert aim 1 table to word
aim3_tbl %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "aim3_table.docx")

```


# Sensitivity Analysis: Housing unaffordability and instability
Sensitivity analysis was conducted to evaluate the sensitivity of results to the distinct dimensions of housing insecurity used in this study
```{r sensitivity}


#unaffordability unadjusted model 1
 affordability_model1<- svyglm(diabetes_num ~ unaffordability,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


#organized table results
 affordability_model1_tbl<-affordability_model1 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = unaffordability,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(unaffordability ~ "Housing unaffordability") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 2: Prevalence Ratios for Diabetes**")  %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")
 
#affordability  main model, model 2 
affordability_model2<- svyglm(diabetes_num ~ unaffordability+agegroup+race+gender+poverty+education+marital,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


#organized table results
 affordability_model2_tbl<-affordability_model2 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = unaffordability,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(unaffordability ~ "Housing unaffordability") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 2: Prevalence Ratios for Diabetes**")  %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


#affordability  model 3
affordability_model3<- svyglm(diabetes_num ~ unaffordability+agegroup+race+gender+poverty+education+marital+bmi+physical_activity,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


#organized table results
 affordability_model3_tbl<-affordability_model3 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = unaffordability,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(unaffordability ~ "Housing unaffordability") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 3: Prevalence Ratios for Diabetes**")  %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")

 
 
 
 affordability_sensitivty_tbl <- tbl_merge(
  tbls = list(affordability_model1_tbl, affordability_model2_tbl, affordability_model3_tbl),
  tab_spanner = c("**Unadjusted**", "**Adjisted**", "**Fully Adjusted**")
)
 
 
 
 
 #Instability unadjusted model 1 
 
  instability_model1<- svyglm(diabetes_num ~ instability,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)


 instability_model1_tbl<-instability_model1 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = instability,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(instability ~ "Housing instability") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 3: Prevalence Ratios for Diabetes**") %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")
 
#instability main model, model 2 
instability_model2<- svyglm(diabetes_num ~ instability+agegroup+race+gender+poverty+education+marital,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)

 instability_model2_tbl<-instability_model2 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = instability,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(instability ~ "Housing instability") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 2: Prevalence Ratios for Diabetes**") %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")

 #instability  model 3
instability_model3<- svyglm(diabetes_num ~ instability+agegroup+race+gender+poverty+education+marital+bmi+physical_activity,
                design = chis_design,
                family = poisson(link = "log"),
                rescale=TRUE)

 instability_model3_tbl<-instability_model3 %>%
  tbl_regression(
    exponentiate = TRUE, 
    include = instability,
    # This changes the "Characteristic" column to "Variable" if you prefer
    label = list(instability ~ "Housing instability") 
  ) %>%
  bold_labels() %>%
  modify_column_hide(p.value) %>%
  # This line changes the column header from IRR to PR
  modify_header(estimate ~ "**PR**") %>% 
  modify_caption("**Model 2: Prevalence Ratios for Diabetes**") %>%
  modify_footnote(estimate ~ "PR = Prevalence Ratio; CI = Confidence Interval")


 
  instability_sensitivty_tbl<-tbl_merge(
  tbls = list(instability_model1_tbl, instability_model2_tbl, instability_model3_tbl),
  tab_spanner = c("**Unadjusted**", "**Adjusted**", "**Fully Adjusted**")
)
  

  

  
#merge instability and affordability PR tables
 
 ## stack all three outcomes into one table
sensitivtyaim_tbl<-tbl_stack(
  tbls = list(instability_sensitivty_tbl, affordability_sensitivty_tbl),
  group_header = c("Instable Housing", "Unaffordable  Housing")
) %>%
  modify_caption("**Sensitivity Analysis**")

sensitivtyaim_tbl

#convert aim 1 table to word
sensitivtyaim_tbl %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "sensitivity_table.docx")

```

# Supplementary section

## Unweighted Table 1

```{r sup}

#unweighted table 1 


tbl1_unweighted <- tbl_summary(
    newchis,
    include = c(gender, agegroup, race, education, urban, poverty, marital, diabetes, bmi, physical_activity, generalhealth, hypertension, heartdisease),
    by = housing_insecure,
    type = list(gender ~ "dichotomous",
                urban ~ "dichotomous",
                marital ~ "dichotomous"),
    value = list(gender ~ "Male",
                 urban ~ "Urban",
                 marital ~ "partnered"),
    label = list(gender ~ "Male",
                 urban ~ "Urban Residence",
                 agegroup ~ "Age Group",
                 race ~ "Race/Ethnicity",
                 education ~ "Education",
                 poverty ~ "Poverty Level",
                 marital ~ "Married/Living with partner",
                 diabetes ~ "Diabetes yes",
                 bmi ~ "BMI",
                 physical_activity ~ "Physical Activity yes",
                 generalhealth ~ "Self reported health",
                 hypertension ~ "Hypertension",
                 heartdisease ~ "Heart disease"
                ),
    # FIX HERE: Change {p} to {n} ({p})
    statistic = list(all_categorical() ~ "{n} ({p})"), 
    # This applies 0 decimal places to the count (n) and 1 decimal place to the percent (p)
    digits = list(all_categorical() ~ c(0, 1))        
  ) %>%
  add_overall() %>%
  modify_table_body(~ .x %>% relocate(stat_2, .before = stat_1)) %>%
  modify_spanning_header(all_stat_cols() ~ "**Housing Insecurity Status**")%>%
  bold_labels()


#save table as word doc
library(flextable)
tbl1_unweighted %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "tbl1_unweighted.docx")

```


## Weighted table 1 restricted by race/ethnicity
```{r supplemnetary}

#Hispanic only
tbl1_weighted_hispanic <- subset(chis_design1, race == "Hispanic") %>% # Adjust string to match your data
  tbl_svysummary(
    include = c(gender, agegroup, education, urban, poverty, marital, diabetes, bmi, physical_activity, generalhealth, hypertension, heartdisease), # Removed 'race' from include
    by = housing_insecure,
    type = list(gender ~ "dichotomous",
                urban ~ "dichotomous",
                marital ~ "dichotomous"),
    value = list(gender ~ "Male",
                 urban ~ "Urban",
                 marital ~ "partnered"),
    label = list(gender ~ "Male",
                 urban ~ "Urban Residence",
                 agegroup ~ "Age Group",
                 education ~ "Education",
                 poverty ~ "Poverty Level",
                 marital ~ "Married/Living with partner",
                 diabetes ~ "Diabetes yes",
                 bmi ~ "BMI",
                 physical_activity ~ "Physical Activity yes",
                 generalhealth ~ "Self reported health",
                 hypertension ~ "Hypertension",
                 heartdisease ~ "Heart disease"
                ),
    statistic = list(all_categorical() ~ "{p}"), 
    digits = list(all_categorical() ~ 1)        
  ) %>%
  add_overall() %>%
  modify_table_body(~ .x %>% relocate(stat_2, .before = stat_1)) %>%
  modify_spanning_header(all_stat_cols() ~ "**Housing Insecurity Status (Hispanic Adults)**") %>%
  bold_labels() 

tbl1_weighted_hispanic %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "tbl1_weighted_hispanic.docx")



tbl1_weighted_other <- subset(chis_design1, race == "Other") %>% # Adjust string to match your data
  tbl_svysummary(
    include = c(gender, agegroup, education, urban, poverty, marital, diabetes, bmi, physical_activity, generalhealth, hypertension, heartdisease), # Removed 'race' from include
    by = housing_insecure,
    type = list(gender ~ "dichotomous",
                urban ~ "dichotomous",
                marital ~ "dichotomous"),
    value = list(gender ~ "Male",
                 urban ~ "Urban",
                 marital ~ "partnered"),
    label = list(gender ~ "Male",
                 urban ~ "Urban Residence",
                 agegroup ~ "Age Group",
                 education ~ "Education",
                 poverty ~ "Poverty Level",
                 marital ~ "Married/Living with partner",
                 diabetes ~ "Diabetes yes",
                 bmi ~ "BMI",
                 physical_activity ~ "Physical Activity yes",
                 generalhealth ~ "Self reported health",
                 hypertension ~ "Hypertension",
                 heartdisease ~ "Heart disease"
                ),
    statistic = list(all_categorical() ~ "{p}"), 
    digits = list(all_categorical() ~ 1)        
  ) %>%
  add_overall() %>%
  modify_table_body(~ .x %>% relocate(stat_2, .before = stat_1)) %>%
  modify_spanning_header(all_stat_cols() ~ "**Housing Insecurity Status (Other race Adults)**") %>%
  bold_labels() 

tbl1_weighted_other %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "tbl1_weighted_other.docx")

#Asian

tbl1_weighted_asian <- subset(chis_design1, race == "Asian") %>% # Adjust string to match your data
  tbl_svysummary(
    include = c(gender, agegroup, education, urban, poverty, marital, diabetes, bmi, physical_activity, generalhealth, hypertension, heartdisease), # Removed 'race' from include
    by = housing_insecure,
    type = list(gender ~ "dichotomous",
                urban ~ "dichotomous",
                marital ~ "dichotomous"),
    value = list(gender ~ "Male",
                 urban ~ "Urban",
                 marital ~ "partnered"),
    label = list(gender ~ "Male",
                 urban ~ "Urban Residence",
                 agegroup ~ "Age Group",
                 education ~ "Education",
                 poverty ~ "Poverty Level",
                 marital ~ "Married/Living with partner",
                 diabetes ~ "Diabetes yes",
                 bmi ~ "BMI",
                 physical_activity ~ "Physical Activity yes",
                 generalhealth ~ "Self reported health",
                 hypertension ~ "Hypertension",
                 heartdisease ~ "Heart disease"
                ),
    statistic = list(all_categorical() ~ "{p}"), 
    digits = list(all_categorical() ~ 1)        
  ) %>%
  add_overall() %>%
  modify_table_body(~ .x %>% relocate(stat_2, .before = stat_1)) %>%
  modify_spanning_header(all_stat_cols() ~ "**Housing Insecurity Status (Hispanic Adults)**") %>%
  bold_labels() 

tbl1_weighted_asian %>%
  as_flex_table() %>%
  flextable::save_as_docx(path = "tbl1_weighted_asian.docx")


```




